---
title: "nls_testing_for_diel-variation-in-dive-depth"
author: "David Green"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(tidyverse)
library(robustbase)
```

##Step 1: Read in dive data. Working with the dive_metrics dataset because it has been prefiltered (see seaTracks for filtering procedures). 

As a first stab I am only reading in id, date, max_dep, lon, lat

```{r}

dive <- fread("../../seaTracks/processed_datasets/dive_metrics.csv",
              select = c("id", "date", "max_dep", "lon", "lat")) %>% 
  drop_na(id)
  # filter(id %in% unique(id)[1:5]) # starting with a small subset to get the workflow going
  # filter(id %in% "ct160-336-20") # starting with a small subset to get the workflow going

```

Because we are working with diel changes in dive depth, we need to have a diel characterisation of each dive

```{r}
source("../../Scripts/Scripts from others/From Sophie/DNid_update.R") # bring in function for calculating diel info

dive <- DNid(dive) # run through the function

```

Ok, now that we have an indicator of hours to midnight, we'll try to fit a nonlinear regression model to estimate nighttime, daytime and midday depths

```{r}
dive <- dive %>% # create day column for fitting individual nls models
  mutate(t_c = noon - abs(difftime(date, noon, units = "hours"))) %>%
  # mutate(t_c = difftime(noon, abs(difftime(date, noon, units = "hours")))) %>% 
  mutate(day = as.POSIXct(substr(date, 1, 10), tz = "UTC")) %>% 
  group_by(id) %>% 
  right_join(data.frame(day = seq.POSIXt(min(.$day),
                                         max(.$day),
                                         by = "day"),
                        id = .$id) %>% 
               mutate(n_day = row_number(.))
             )

dive %>% 
  mutate(day = as.POSIXct(day, tz = "UTC")) %>% 
  ggplot(aes(x = day,
             y = max_dep)) + 
  geom_point()

```

Creating an example test, also used to create some template outputs

```{r}


  
# plot(dive$day, dive$max_dep)
# trying with one individual, and for one day
dive_sub <- dive %>% 
  # filter(id %in% unique(id)[4]) %>% 
  filter(id %in% "ct160-336-20") %>% 
  filter(day %in% unique(day)[70:70])

# {par(mfrow = c(1,2))
# plot(dive_sub$hrs_to_noon, dive_sub$max_dep)
# plot(dive_sub$sol_pos, dive_sub$max_dep)
# }
# plot(dive_sub$date, dive_sub$max_dep)
# fit <- nlrob(max_dep ~ hrs_to_midnight, data = dive_sub)

# fit <- nlrob(max_dep ~ (D_n + (D_d - D_n))/(1-exp(as.numeric(difftime(t_m,t_c, units = "hours"))/lamd)),
#              data = dive_sub,
#              start = c(D_n = 100, D_d = 500, t_m = ))

# Starting depths

if(any(dive_sub$diel == "night")){ # if night dives available
  D_n <- median(dive_sub$max_dep[dive_sub$diel == "night"], na.rm = T)
}else{
    D_n <- median(dive_sub$max_dep[dive_sub$diel %in% c("dawn", "dusk")], na.rm = T) # if only twilight dives available
  }

if(any(dive_sub$diel == "day")){ # if day dives available
  D_d <- median(dive_sub$max_dep[dive_sub$diel == "day"], na.rm = T)
}else{
    D_d <- median(dive_sub$max_dep[dive_sub$diel %in% c("dawn", "dusk")], na.rm = T) # if only twilight dives available
  }

# D_n <- median(dive_sub$max_dep[dive_sub$diel == "night"], na.rm = T)
# D_d <- median(dive_sub$max_dep[dive_sub$diel == "day"], na.rm = T)

# Try with distance from noon
# 
# t_m <- mean(range(dive_sub$hrs_to_noon)) # inflection point for slope (hours to noon)
# lamd_noon <- cor(dive_sub$max_dep, dive_sub$hrs_to_noon)
# 
# fit_noon <- nlrob(max_dep ~ D_n + (D_d - D_n)/(1+exp((t_m - hrs_to_noon)/lamd_noon)),
#              data = dive_sub,
#              start = c(D_n = D_n, 
#                        D_d = D_d, 
#                        t_m = t_m, 
#                        lamd_noon = lamd_noon),
#              # lower = c(D_n = 0,
#              #           D_d = 0),
#              algorithm = "port")
# summary(fit_noon)
# Try with solar position instead

sol_mid <- mean(range(dive_sub$sol_pos, na.rm = T)) # inflection point for slope (solar position)
lamd_sol <- cor(dive_sub$max_dep, dive_sub$sol_pos)

fit_sol <- nlrob(max_dep ~ D_n + (D_d - D_n)/(1+exp((sol_mid - sol_pos)/lamd_sol)),
             data = dive_sub,
             start = c(D_n = D_n, 
                       D_d = D_d, 
                       sol_mid = sol_mid, 
                       lamd_sol = lamd_sol),
             algorithm = "port",
             maxit = 500)

fit_dat <- as.data.frame(summary(fit_sol)$coefficients) %>% 
  rownames_to_column() %>% 
  pivot_wider(names_from = rowname,
              values_from = -rowname)

# summary(fit_sol)


# Creating generic output if test fails to converge
out_fail <- fit_dat %>% 
  mutate(across(everything(), ~ NA_real_)) %>% 
  mutate(id = NA_character_,
         day_s = NA_POSIXct_,
         day_e = NA_POSIXct_
         )

```

Now, let's try iterating across the whole dataset for a single individual
```{r}

fits_sum <- out_fail[0,]
wind_size <- 3

for(jj in  unique(dive$id)){
  
  dive_sub <- dive %>% 
    filter(id %in% jj)

  for(ii in 1:(max(unique(dive_sub$n_day)) - (wind_size - 1))){
    cat("\r testing ",jj, "| day number: ", ii, "                          ")
    
    rm(D_n, D_d, sol_mid, lamd_sol)
    
    test_dat <- dive_sub %>% 
      # filter(n_day >= ii & n_day <= (ii+1)) %>% 
      filter(n_day >= ii & n_day <= (ii + (wind_size - 1))) %>% 
      drop_na(diel)
    
    suppressWarnings({
      # Select starting parameter values
      # Night time depth
      if(any(test_dat$diel == "night")){ # if night dives available
      D_n <- mean(test_dat$max_dep[test_dat$diel == "night"], na.rm = T)
      }else{
        D_n <- mean(test_dat$max_dep[test_dat$diel %in% c("dawn", "dusk")], na.rm = T) # if only twilight dives available
      }
    
    # Day time depth
      if(any(test_dat$diel == "day")){ # if day dives available
        D_d <- mean(test_dat$max_dep[test_dat$diel == "day"], na.rm = T)
      }else{
          D_d <- mean(test_dat$max_dep[test_dat$diel %in% c("dawn", "dusk")], na.rm = T) # if only twilight dives available
        }
      
      # inflection point for slope (solar position)
      sol_mid <- mean(range(test_dat$sol_pos, na.rm = T))
      
       # slope of logicstic curve
      lamd_sol <- cor(test_dat$max_dep, test_dat$sol_pos)
        
      
      
      
      
      # fit robust nls model
      suppressWarnings(
      fit_sol <- tryCatch(
        expr = {
          nlrob(max_dep ~ D_n + (D_d - D_n)/(1+exp((sol_mid - sol_pos)/lamd_sol)),
                 data = test_dat,
                 start = c(D_n = D_n, 
                           D_d = D_d, 
                           sol_mid = sol_mid, 
                           lamd_sol = lamd_sol)
                )
        },
        error = function(e){
          return(NULL)
        }
      )
      )
      
      # extract summary output
      if(!is.null(fit_sol)){
        fit_dat <- as.data.frame(summary(fit_sol)$coefficients) %>% 
          rownames_to_column() %>% 
          pivot_wider(names_from = rowname,
                      values_from = -rowname) %>% 
          mutate(id = test_dat$id[1],
                 day_s = min(test_dat$day),
                 day_e = max(test_dat$day),
                 D_diff = Estimate_D_d - Estimate_D_n
                 )
      }else{
        fit_dat <- out_fail %>% 
          mutate(id = test_dat$id[1],
                 day_s = min(test_dat$day),
                 day_e = max(test_dat$day),
                 D_d_start = D_d,
                 D_n_start = D_n,
                 D_diff_start = D_d - D_n
                 )
      }
      
      # append to summary dataset
      fits_sum <- bind_rows(fits_sum, fit_dat)
    })
  }
}

```





